---
title: "Great Bear Rainforest Salmon Report - Supplemental Information"
author: "Coastland"
date: "2024-02-03"
output:
  bookdown::gitbook: default
  html_document:
    df_print: paged
---


# index

This webpage hosts summary plots for Rosenberger and Bottoms (_in prep_). Analyses use data from DFO's New Salmon Escapement Database Systems (NUSEDs) up to 2022. 

This supplemental information report breaks down trends in salmon escapements by species, system, and conservation unit. Specifics are defined at the head of each section found in the navigation panel. 

# Escapement plots

Escapement plots over time, by species, by Conservation Unit. Colours represent whether or not the stream qualified for analysis in this study. Solid vertical lines denote the establishment of the Great Bear Rainforest Special Protection Area (GBR SPA) in 2007 and dashed vertical lines represent the cutoff for the current generation included in the analyses. 


```{r Define custom data for escapement bubble plots, include = FALSE, message = FALSE}

library(tidyverse)
library(kableExtra)

# Generation length lkp
gen.len <- data.frame(SPECIES_QUALIFIED = c('CK', 'SEL', 'CM', 
                                            'CO', 'PKE', 'PKO'),
                      gen.len = c(4, 4, 4, 3, 2, 2))

# Periods lkp
periods <- gen.len %>%
  mutate(cyear = 2022,
         gbr.est = 2007,
         cgen.y1 = ifelse(SPECIES_QUALIFIED == 'PKE' | SPECIES_QUALIFIED == 'PKO', 
                        cyear - gen.len - 1,  # so PKE/PKO have 2 return years
                        cyear - gen.len + 1), 
         cgen.y2 = cyear,
         pregbr.y1 = gbr.est - (3 * gen.len) + 1,
         pregbr.y2 = gbr.est,
         postgbr.y1 = gbr.est,
         postgbr.y2 = gbr.est + (3 * gen.len) - 1,
         all.y1 = 1920, # earliest year in nuseds
         all.y2 = gbr.est - 1)

gbr.filt <- readRDS('data/gbr.filt.rds')


link <- read.csv('data/stream id to analyse link.csv')
data <- readRDS("data/nuseds.rds") %>%
  merge(link) %>%
  merge(periods, by = "SPECIES_QUALIFIED") %>%
  filter(WATERBODY %in% gbr.filt$WATERBODY) %>%
  mutate(year = as.numeric(year))

dat <- readRDS('data/dat.rds')
ratios <- readRDS('data/ratios.rds')

```



## Summary Table

```{r Escapement summary table, echo = FALSE}

kable(
  data %>%
  select(SPECIES_QUALIFIED, SPECIES, WATERBODY, ANALYSE) %>%
  unique() %>%
  mutate(SPECIES = ifelse(SPECIES_QUALIFIED == 'PKE', "Pink (Even years)", SPECIES),
         SPECIES = ifelse(SPECIES_QUALIFIED == 'PKO', "Pink (Odd years)", SPECIES)) %>%
  count(SPECIES, ANALYSE) %>%
  rename(Species = SPECIES,
         `Included in analysis` = ANALYSE,
         `Number of streams` = n) %>%
  arrange(desc(`Included in analysis`))
)


```


## Chinook

```{r Chinook escapement bubble plots, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 22, fig.width=10}

plot.data<-data%>%filter(SPECIES_QUALIFIED=='CK')
g.len <- unique(plot.data$cgen.y1)
  
ggplot(plot.data,aes(x=year,y=WATERBODY,size=escapement, color = ANALYSE))+
  geom_point(alpha=.5)+
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  labs(y="Stream",x="Year", subtitle = 'Chinook', color = "Analysed?",
       size = "Escapement")+
  theme_bw()+
  geom_vline(aes(xintercept = 2007)) +
  geom_vline(aes(xintercept = g.len), linetype = 2) +
  theme(strip.text.y = element_text(angle=0,size=8),
        axis.text.x=element_text(angle = 0, size=12)) + 
  scale_color_brewer(palette = 'Set1')

```


## Chum
```{r Chum escapement bubble plots, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 65, fig.width=10}

plot.data<-data%>%filter(SPECIES_QUALIFIED=='CM')
g.len <- unique(plot.data$cgen.y1)
  
ggplot(plot.data,aes(x=year,y=WATERBODY,size=escapement, color = ANALYSE))+
  geom_point(alpha=.5)+
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  labs(y="Stream",x="Year", subtitle = 'Chum', color = "Analysed?",
       size = "Escapement")+
  theme_bw()+
  geom_vline(aes(xintercept = 2007)) +
  geom_vline(aes(xintercept = g.len), linetype = 2) +
  theme(strip.text.y = element_text(angle=0,size=8),
        axis.text.x=element_text(angle = 0, size=12)) + 
  scale_color_brewer(palette = 'Set1')

```


## Coho
```{r Coho escapement bubble plots, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 65, fig.width=10}

plot.data<-data%>%filter(SPECIES_QUALIFIED=='CO')
g.len <- unique(plot.data$cgen.y1)
  
ggplot(plot.data,aes(x=year,y=WATERBODY,size=escapement, color = ANALYSE))+
  geom_point(alpha=.5)+
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  labs(y="Stream",x="Year", subtitle = 'Coho', color = "Analysed?",
       size = "Escapement")+
  theme_bw()+
  geom_vline(aes(xintercept = 2007)) +
  geom_vline(aes(xintercept = g.len), linetype = 2) +
  theme(strip.text.y = element_text(angle=0,size=8),
        axis.text.x=element_text(angle = 0, size=12)) + 
  scale_color_brewer(palette = 'Set1')

```

## Pink (Even years)
```{r Pink (even years) escapement bubble plots, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 65, fig.width=10}

plot.data<-data%>%filter(SPECIES_QUALIFIED=='PKE')
g.len <- unique(plot.data$cgen.y1)
  
ggplot(plot.data,aes(x=year,y=WATERBODY,size=escapement, color = ANALYSE))+
  geom_point(alpha=.5)+
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  labs(y="Stream",x="Year", subtitle = 'Pink (Even Years)', color = "Analysed?",
       size = "Escapement")+
  theme_bw()+
  geom_vline(aes(xintercept = 2007)) +
  geom_vline(aes(xintercept = g.len), linetype = 2) +
  theme(strip.text.y = element_text(angle=0,size=8),
        axis.text.x=element_text(angle = 0, size=12)) + 
  scale_color_brewer(palette = 'Set1')

```


## Pink (odd years)
```{r Pink (odd years) escapement bubble plots, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 65, fig.width=10}

plot.data<-data%>%filter(SPECIES_QUALIFIED=='PKO')
g.len <- unique(plot.data$cgen.y1)
  
ggplot(plot.data,aes(x=year,y=WATERBODY,size=escapement, color = ANALYSE))+
  geom_point(alpha=.5)+
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  labs(y="Stream",x="Year", subtitle = 'Pink (Odd Years)', color = "Analysed?",
       size = "Escapement")+
  theme_bw()+
  geom_vline(aes(xintercept = 2007)) +
  geom_vline(aes(xintercept = g.len), linetype = 2) +
  theme(strip.text.y = element_text(angle=0,size=8),
        axis.text.x=element_text(angle = 0, size=12)) + 
  scale_color_brewer(palette = 'Set1')

```



## Sockeye
```{r Sockeye escapement bubble plots, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 26, fig.width=10}

plot.data<-data%>%filter(SPECIES_QUALIFIED=='SEL')
g.len <- unique(plot.data$cgen.y1)
  
ggplot(plot.data,aes(x=year,y=WATERBODY,size=escapement, color = ANALYSE))+
  geom_point(alpha=.5)+
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  labs(y="Stream",x="Year", subtitle = 'Sockeye (Lake)', color = "Analysed?",
       size = "Escapement")+
  theme_bw()+
  geom_vline(aes(xintercept = 2007)) +
  geom_vline(aes(xintercept = g.len), linetype = 2) +
  theme(strip.text.y = element_text(angle=0,size=8),
        axis.text.x=element_text(angle = 0, size=12)) + 
  scale_color_brewer(palette = 'Set1')

```

# Trends

This section compares trends in escapements by Conservation Unit on all waterbodies selected for analysis by period.

Periods are defined as: 

`all` : 

- Baseline metric, represents the trend over all years from 1920 - 2006 (one year before the establishment of the GBR SPA)

`cgen` : 

- Trend for the current generation (2022 minus one generation length)

`preGBR` :

- Comparison metric representing the trend for three generations before the establishment of the GBR SPA

`postGBR` :

- Comparison metric representing the trend for three generations after the establishment of the GBR SPA


## All species

Comparisons across all Conservation Units

```{r Trends by all species by period, echo = FALSE, message = FALSE,warning = FALSE, fig.width=8}

ggplot(dat, aes(x = spp.qual, y = trend, fill = period)) +
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
  theme_bw() + 
  geom_hline(yintercept = 0) +
  labs(x = "Species", y = "Trend over period", fill = "Period") + 
  scale_fill_brewer(palette = 'Set1')

```


## Chinook
```{r Chinook trends by period by CU, echo = FALSE, message = FALSE,warning = FALSE, fig.height= 12, fig.width= 8}

ggplot(dat %>% filter(spp.qual == 'CK'),
       aes(x = period, y = trend, fill = period)) + 
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
  theme_bw() +
  geom_hline(yintercept = 0) +
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  theme(strip.text.y = element_text(angle=0,size=8)) +
  scale_fill_brewer(palette = "Set1") + 
  coord_flip() +
  labs(x = "Conservation Unit by Period", y = "Trend") +
  theme(legend.position = 'top', axis.text.x=element_text(angle = 0)) + 
  scale_fill_brewer(palette = 'Set1')


```


## Chum
```{r Chum trends by period by CU, echo = FALSE, message = FALSE,warning = FALSE, fig.height= 12, fig.width= 8}

ggplot(dat %>% filter(spp.qual == 'CM'),
       aes(x = period, y = trend, fill = period)) + 
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
  theme_bw() +
  geom_hline(yintercept = 0) +
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  theme(strip.text.y = element_text(angle=0,size=8)) +
  scale_fill_brewer(palette = "Set1") + 
  coord_flip() +
  labs(x = "Conservation Unit by Period", y = "Trend") +
  theme(legend.position = 'top', axis.text.x=element_text(angle = 0)) + 
  scale_fill_brewer(palette = 'Set1')


```

## Coho
```{r Coho trends by period by CU, echo = FALSE, message = FALSE,warning = FALSE, fig.height= 12, fig.width= 8}

ggplot(dat %>% filter(spp.qual == 'CO'),
       aes(x = period, y = trend, fill = period)) + 
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
  theme_bw() +
  geom_hline(yintercept = 0) +
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  theme(strip.text.y = element_text(angle=0,size=8)) +
  scale_fill_brewer(palette = "Set1") + 
  coord_flip() +
  labs(x = "Conservation Unit by Period", y = "Trend") +
  theme(legend.position = 'top', axis.text.x=element_text(angle = 0)) + 
  scale_fill_brewer(palette = 'Set1')


```

## Pink (Even years)
```{r Pink (Even years) trends by period by CU, echo = FALSE, message = FALSE,warning = FALSE, fig.height= 5, fig.width= 8}

ggplot(dat %>% filter(spp.qual == 'PKE'),
       aes(x = period, y = trend, fill = period)) + 
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
  theme_bw() +
  geom_hline(yintercept = 0) +
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  theme(strip.text.y = element_text(angle=0,size=8)) +
  scale_fill_brewer(palette = "Set1") + 
  coord_flip() +
  labs(x = "Conservation Unit by Period", y = "Trend") +
  theme(legend.position = 'top', axis.text.x=element_text(angle = 0)) + 
  scale_fill_brewer(palette = 'Set1')


```

## Pink (Odd years)
```{r Pink (Odd years) trends by period by CU, echo = FALSE, message = FALSE,warning = FALSE, fig.height= 8, fig.width= 8}

ggplot(dat %>% filter(spp.qual == 'PKO'),
       aes(x = period, y = trend, fill = period)) + 
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
  theme_bw() +
  geom_hline(yintercept = 0) +
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  theme(strip.text.y = element_text(angle=0,size=8)) +
  scale_fill_brewer(palette = "Set1") + 
  coord_flip() +
  labs(x = "Conservation Unit by Period", y = "Trend") +
  theme(legend.position = 'top', axis.text.x=element_text(angle = 0)) + 
  scale_fill_brewer(palette = 'Set1')


```

## Sockeye
```{r Sockeye trends by period by CU, echo = FALSE, message = FALSE,warning = FALSE, fig.height= 20, fig.width= 8}

ggplot(dat %>% filter(spp.qual == 'SEL'),
       aes(x = period, y = trend, fill = period)) + 
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
  theme_bw() +
  geom_hline(yintercept = 0) +
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  theme(strip.text.y = element_text(angle=0,size=8)) +
  scale_fill_brewer(palette = "Set1") + 
  coord_flip() +
  labs(x = "Conservation Unit by Period", y = "Trend") +
  theme(legend.position = 'top', axis.text.x=element_text(angle = 0)) + 
  scale_fill_brewer(palette = 'Set1')


```

# Trend Ratios

This section shows how trends compare across periods. Comparison periods are defined as:

`trnd.cgen.all`:

- Ratio comparing how the escapement trend of the current generation is performing against the baseline trend over all years prior to the GBR SPA.

`trnd.cgen.pre`:

- Ratio comparing how the escapement trend of the current generation `cgen` is performing against the trend over three generations prior to the GBR SPA.

`trnd.post.all`:

- Ratio comparing how the escapement trend of the three generations following the GBR SPA is performing against the baseline trend over all years prior to the GBR SPA.

`trnd.post.pre`:

- Ratio comparing how the escapement trend of the three generations following the GBR SPA is performing against the three generations prior to the GBR SPA.

## All species

Comparisons across all Conservation Units

```{r Trend ratios all species all CU, echo = FALSE, warning = FALSE, message = FALSE, fig.width=8}

ggplot(ratios, aes(x = spp.qual, y = trend.ratio, fill = trend.ratio.period)) + 
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) + 
  theme_bw() +
  geom_hline(yintercept = 0) +
  scale_fill_brewer(palette = "Set1") + 
  # coord_flip() + 
  labs(x = "Species", y = "Ratio", 
       fill = "Comparison period")

```


## Chinook
```{r Chinook ratio by CU, echo = FALSE, message = FALSE, warning=FALSE,, fig.height=8, fig.width=8}

  ggplot(ratios %>% filter(spp.qual == 'CK'), aes(x = trend.ratio.period, y = trend.ratio,
                                                 fill = trend.ratio.period)) + 
    geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
    theme_bw() +
    facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
    theme(strip.text.y = element_text(angle=0,size=8)) +
    geom_hline(yintercept = 0) +
    scale_fill_brewer(palette = "Set1") + 
    coord_flip() + 
    labs(x = "Conservation Unit by Comparison Period", y = "Trend ratio", fill = "Comparison period") +
    theme(legend.position = 'top')

```


## Chum
```{r Chum ratio by CU, echo = FALSE, message = FALSE, warning=FALSE,, fig.height=11, fig.width=8}

  ggplot(ratios %>% filter(spp.qual == 'CM'), aes(x = trend.ratio.period, y = trend.ratio,
                                                 fill = trend.ratio.period)) + 
    geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
    theme_bw() +
    facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
    theme(strip.text.y = element_text(angle=0,size=8)) +
    geom_hline(yintercept = 0) +
    scale_fill_brewer(palette = "Set1") + 
    coord_flip() + 
    labs(x = "Conservation Unit by Comparison Period", y = "Trend ratio", fill = "Comparison period") +
    theme(legend.position = 'top')

```

## Coho
```{r Coho ratio by CU, echo = FALSE, message = FALSE, warning=FALSE,, fig.height=8, fig.width=8}

  ggplot(ratios %>% filter(spp.qual == 'CO'), aes(x = trend.ratio.period, y = trend.ratio,
                                                 fill = trend.ratio.period)) + 
    geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
    theme_bw() +
    facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
    theme(strip.text.y = element_text(angle=0,size=8)) +
    geom_hline(yintercept = 0) +
    scale_fill_brewer(palette = "Set1") + 
    coord_flip() + 
    labs(x = "Conservation Unit by Comparison Period", y = "Trend ratio", fill = "Comparison period") +
    theme(legend.position = 'top')

```

## Pink (Even years)
```{r Pink (even years) ratio by CU, echo = FALSE, message = FALSE, warning=FALSE,, fig.height=5, fig.width=8}

  ggplot(ratios %>% filter(spp.qual == 'PKE'), aes(x = trend.ratio.period, y = trend.ratio,
                                                 fill = trend.ratio.period)) + 
    geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
    theme_bw() +
    facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
    theme(strip.text.y = element_text(angle=0,size=8)) +
    geom_hline(yintercept = 0) +
    scale_fill_brewer(palette = "Set1") + 
    coord_flip() + 
    labs(x = "Conservation Unit by Comparison Period", y = "Trend ratio", fill = "Comparison period") +
    theme(legend.position = 'top')

```

## Pink (Odd years)
```{r Pink (odd years) ratio by CU, echo = FALSE, message = FALSE, warning=FALSE,, fig.height=8, fig.width=8}

  ggplot(ratios %>% filter(spp.qual == 'PKO'), aes(x = trend.ratio.period, y = trend.ratio,
                                                 fill = trend.ratio.period)) + 
    geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
    theme_bw() +
    facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
    theme(strip.text.y = element_text(angle=0,size=8)) +
    geom_hline(yintercept = 0) +
    scale_fill_brewer(palette = "Set1") + 
    coord_flip() + 
    labs(x = "Conservation Unit by Comparison Period", y = "Trend ratio", fill = "Comparison period") +
    theme(legend.position = 'top')

```

## Sockeye
```{r Sockeye ratio by CU, echo = FALSE, message = FALSE, warning=FALSE,, fig.height=10, fig.width=8}

  ggplot(ratios %>% filter(spp.qual == 'SEL'), aes(x = trend.ratio.period, y = trend.ratio,
                                                 fill = trend.ratio.period)) + 
    geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
    theme_bw() +
    facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
    theme(strip.text.y = element_text(angle=0,size=8)) +
    geom_hline(yintercept = 0) +
    scale_fill_brewer(palette = "Set1") + 
    coord_flip() + 
    labs(x = "Conservation Unit by Comparison Period", y = "Trend ratio", fill = "Comparison period") +
    theme(legend.position = 'top')

```

# Geometric means

This section compares the geometric means of escapements by Conservation Unit on all waterbodies selected for analysis by period.

Periods are defined as: 

`all` : 

- Baseline metric, represents the geomean over all years from 1920 - 2006 (one year before the establishment of the GBR SPA)

`cgen` : 

- Geomean for the current generation (2022 minus one generation length)

`preGBR` :

- Comparison metric representing the geomean for three generations before the establishment of the GBR SPA

`postGBR` :

- Comparison metric representing the geomean for three generations after the establishment of the GBR SPA

## All species

Comparisons across all Conservation Units

```{r Geomeans by all species by period, echo = FALSE, message = FALSE,warning = FALSE, fig.width=8}

ggplot(dat, aes(x = spp.qual, y = (log(gm) + 1), fill = period)) +
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
  theme_bw() + 
  labs(x = "Species", y = "log(Geomean) + 1", fill = "Period") + 
  scale_fill_brewer(palette = 'Set1') 

```

## Chinook
```{r Chinook geomeans by period by CU, echo = FALSE, message = FALSE,warning = FALSE, fig.height= 12, fig.width= 8}

ggplot(dat %>% filter(spp.qual == 'CK'),
       aes(x = period, y = (log(gm) + 1), fill = period)) + 
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
  theme_bw() +
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  theme(strip.text.y = element_text(angle=0,size=8)) +
  scale_fill_brewer(palette = "Set1") + 
  coord_flip() +
  labs(x = "Conservation Unit by Period", y = "log(Geomean) + 1") +
  theme(legend.position = 'top', axis.text.x=element_text(angle = 0)) + 
  scale_fill_brewer(palette = 'Set1')


```

## Chum
```{r Chum geomeans by period by CU, echo = FALSE, message = FALSE,warning = FALSE, fig.height= 12, fig.width= 8}

ggplot(dat %>% filter(spp.qual == 'CM'),
       aes(x = period, y = (log(gm) + 1), fill = period)) + 
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
  theme_bw() +
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  theme(strip.text.y = element_text(angle=0,size=8)) +
  scale_fill_brewer(palette = "Set1") + 
  coord_flip() +
  labs(x = "Conservation Unit by Period", y = "log(Geomean) + 1") +
  theme(legend.position = 'top', axis.text.x=element_text(angle = 0)) + 
  scale_fill_brewer(palette = 'Set1')


```

## Coho
```{r Coho geomeans by period by CU, echo = FALSE, message = FALSE,warning = FALSE, fig.height= 12, fig.width= 8}

ggplot(dat %>% filter(spp.qual == 'CO'),
       aes(x = period, y = (log(gm) + 1), fill = period)) + 
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
  theme_bw() +
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  theme(strip.text.y = element_text(angle=0,size=8)) +
  scale_fill_brewer(palette = "Set1") + 
  coord_flip() +
  labs(x = "Conservation Unit by Period", y = "log(Geomean) + 1") +
  theme(legend.position = 'top', axis.text.x=element_text(angle = 0)) + 
  scale_fill_brewer(palette = 'Set1')


```

## Pink (Even years)
```{r Pink (Even years) geomeans by period by CU, echo = FALSE, message = FALSE,warning = FALSE, fig.height= 5, fig.width= 8}

ggplot(dat %>% filter(spp.qual == 'PKE'),
       aes(x = period, y = (log(gm) + 1), fill = period)) + 
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
  theme_bw() +
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  theme(strip.text.y = element_text(angle=0,size=8)) +
  scale_fill_brewer(palette = "Set1") + 
  coord_flip() +
  labs(x = "Conservation Unit by Period", y = "log(Geomean) + 1") +
  theme(legend.position = 'top', axis.text.x=element_text(angle = 0)) + 
  scale_fill_brewer(palette = 'Set1')


```


## Pink (Odd years)
```{r Pink (odd years) geomeans by period by CU, echo = FALSE, message = FALSE,warning = FALSE, fig.height= 8, fig.width= 8}

ggplot(dat %>% filter(spp.qual == 'PKO'),
       aes(x = period, y = (log(gm) + 1), fill = period)) + 
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
  theme_bw() +
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  theme(strip.text.y = element_text(angle=0,size=8)) +
  scale_fill_brewer(palette = "Set1") + 
  coord_flip() +
  labs(x = "Conservation Unit by Period", y = "log(Geomean) + 1") +
  theme(legend.position = 'top', axis.text.x=element_text(angle = 0)) + 
  scale_fill_brewer(palette = 'Set1')


```

## Sockeye
```{r Sockeye geomeans by period by CU, echo = FALSE, message = FALSE,warning = FALSE, fig.height= 15, fig.width= 8}

ggplot(dat %>% filter(spp.qual == 'SEL'),
       aes(x = period, y = (log(gm) + 1), fill = period)) + 
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
  theme_bw() +
  facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
  theme(strip.text.y = element_text(angle=0,size=8)) +
  scale_fill_brewer(palette = "Set1") + 
  coord_flip() +
  labs(x = "Conservation Unit by Period", y = "log(Geomean) + 1") +
  theme(legend.position = 'top', axis.text.x=element_text(angle = 0)) + 
  scale_fill_brewer(palette = 'Set1')


```

# Geomean Ratios

This section shows how geometric means compare across periods. Comparison periods are defined as:

`geo.cgen.all`:

- Ratio comparing how the geometric mean of the current generation is performing against the baseline trend over all years prior to the GBR SPA.

`geo.cgen.pre`:

- Ratio comparing how the geometric mean of the current generation `cgen` is performing against the trend over three generations prior to the GBR SPA.

`geo.post.all`:

- Ratio comparing how the geometric mean of the three generations following the GBR SPA is performing against the baseline trend over all years prior to the GBR SPA.

`geo.post.pre`:

- Ratio comparing how the geometric mean of the three generations following the GBR SPA is performing against the three generations prior to the GBR SPA.

## All species

Comparisons across all Conservation Units

```{r Geomean ratios all species all CU, echo = FALSE, warning = FALSE, message = FALSE, fig.width=8}

ggplot(ratios, aes(x = spp.qual, y = geo.ratio, fill = geo.ratio.period)) + 
  geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) + 
  theme_bw() +
  scale_fill_brewer(palette = "Set1") + 
  labs(x = "Species", y = "Geomean ratio", 
       fill = "Comparison period")

```


## Chinook
```{r Chinook geomean ratio by CU, echo = FALSE, message = FALSE, warning=FALSE,, fig.height=8, fig.width=8}

  ggplot(ratios %>% filter(spp.qual == 'CK'), aes(x = geo.ratio.period, y = geo.ratio,
                                                 fill = geo.ratio.period)) + 
    geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
    theme_bw() +
    facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
    theme(strip.text.y = element_text(angle=0,size=8)) +
    scale_fill_brewer(palette = "Set1") + 
    coord_flip() + 
    labs(x = "Conservation Unit by Comparison Period", y = "Geomean ratio", fill = "Comparison period") +
    theme(legend.position = 'top')

```


## Chum
```{r Chum geomean ratio by CU, echo = FALSE, message = FALSE, warning=FALSE,, fig.height=12, fig.width=8}

  ggplot(ratios %>% filter(spp.qual == 'CM'), aes(x = geo.ratio.period, y = geo.ratio,
                                                 fill = geo.ratio.period)) + 
    geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
    theme_bw() +
    facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
    theme(strip.text.y = element_text(angle=0,size=8)) +
    scale_fill_brewer(palette = "Set1") + 
    coord_flip() + 
    labs(x = "Conservation Unit by Comparison Period", y = "Geomean ratio", fill = "Comparison period") +
    theme(legend.position = 'top')

```

## Coho
```{r Coho geomean ratio by CU, echo = FALSE, message = FALSE, warning=FALSE,, fig.height=9, fig.width=8}

  ggplot(ratios %>% filter(spp.qual == 'CO'), aes(x = geo.ratio.period, y = geo.ratio,
                                                 fill = geo.ratio.period)) + 
    geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
    theme_bw() +
    facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
    theme(strip.text.y = element_text(angle=0,size=8)) +
    scale_fill_brewer(palette = "Set1") + 
    coord_flip() + 
    labs(x = "Conservation Unit by Comparison Period", y = "Geomean ratio", fill = "Comparison period") +
    theme(legend.position = 'top')

```

## Pink (Even years)
```{r Pink (Even years) geomean ratio by CU, echo = FALSE, message = FALSE, warning=FALSE,, fig.height=5, fig.width=8}

  ggplot(ratios %>% filter(spp.qual == 'PKE'), aes(x = geo.ratio.period, y = geo.ratio,
                                                 fill = geo.ratio.period)) + 
    geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
    theme_bw() +
    facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
    theme(strip.text.y = element_text(angle=0,size=8)) +
    scale_fill_brewer(palette = "Set1") + 
    coord_flip() + 
    labs(x = "Conservation Unit by Comparison Period", y = "Geomean ratio", fill = "Comparison period") +
    theme(legend.position = 'top')

```

## Pink (Odd years)
```{r Pink (odd years) geomean ratio by CU, echo = FALSE, message = FALSE, warning=FALSE,, fig.height=7, fig.width=8}

  ggplot(ratios %>% filter(spp.qual == 'PKO'), aes(x = geo.ratio.period, y = geo.ratio,
                                                 fill = geo.ratio.period)) + 
    geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
    theme_bw() +
    facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
    theme(strip.text.y = element_text(angle=0,size=8)) +
    scale_fill_brewer(palette = "Set1") + 
    coord_flip() + 
    labs(x = "Conservation Unit by Comparison Period", y = "Geomean ratio", fill = "Comparison period") +
    theme(legend.position = 'top')

```

## Sockeye
```{r Sockeye geomean ratio by CU, echo = FALSE, message = FALSE, warning=FALSE,, fig.height=12, fig.width=8}

  ggplot(ratios %>% filter(spp.qual == 'SEL'), aes(x = geo.ratio.period, y = geo.ratio,
                                                 fill = geo.ratio.period)) + 
    geom_boxplot(alpha = .85, position = position_dodge2(preserve = "single")) +
    theme_bw() +
    facet_grid(CU_NAME~.,space="free_y",scale="free_y",
             labeller = label_wrap_gen(15))+
    theme(strip.text.y = element_text(angle=0,size=8)) +
    scale_fill_brewer(palette = "Set1") + 
    coord_flip() + 
    labs(x = "Conservation Unit by Comparison Period", y = "Geomean ratio", fill = "Comparison period") +
    theme(legend.position = 'top')

```